### 2014-12-13 paper.tex
Add to compactness metrics discussion

  Of the efficiency gap computations, the seat share computation appeared to actually have a stronger relationship to compactness, in spite of being a simplified metric.  However, due to the small sample size and the fact that the two metrics produce similar results in many states, this may not be a meaningful difference.  Of the compactness metrics, moment of area performed the best; convex hull area and moment of population also did well.  The area-perimeter metric was ineffective; it did correlate with the seat share, but less strongly, and there were several states with low district compactness with little partisan bias.  The convex hull population metric also did poorly; we conjecture that this is due to the fact that many suburban districts, even those which are reasonably shaped, include much of the nearby city in their convex hull.

@@

  \section{Discussion}
  \subsection{Usefulness of Different Metrics}
  \subsubsection{Compactness Metrics}
  %TODO(benkraft): write the other half of this section, maybe reorganize a bit
  Metrics that use the convex hull of a district are based on the observation that problematic districts are likely to have a section ``cut out'', jut out into a neighboring district, or stretch out in a long, narrow, curved line. Like the perimeter metric, they disadvantage districts with jagged boundaries, although less heavily. The purely geometric version, like other compactness metrics, penalizes states that are non-compact. It particularly affects states that are on a coast and have islands, because a large offshore area gets counted in the convex hull. For example, California's 14th district gets an especially bad score because it combines a small, densely populated area with a group of islands several miles offshore. However, the population-based version does not have these issues. In cases where a state has complex boundaries with other states, the population outside of the state can simply be ignored. In cases where a state has islands, the offshore area has no population so it has no effect.

@@

  %TODO(benkraft): write this section

>>>>

  Of the efficiency gap computations, the seat share computation appeared to actually have a stronger relationship to compactness, in spite of being a simplified metric.  However, due to the small sample size and the fact that the two metrics produce similar results in many states, this may not be a meaningful difference.  Of the compactness metrics, moment of area performed the best; convex hull area and moment of population also did well.  The area-perimeter metric was ineffective; it did correlate with the seat share, but less strongly, and there were several states with low district compactness with little partisan bias.  The convex hull population metric also did less well than the others.

@@

  \section{Discussion}
  \subsection{Usefulness of Different Metrics}
  \subsubsection{Compactness Metrics}
  As measures for gerrymandering, none of the measures of compactness alone is a perfect measure.  In each case, there were some states that scored poorly on a compactness metric, but for which there was no evidence of partisan gerrymandering.  It is possible that in these cases there was a bipartisan gerrymander, or a gerrymander which did not happen to cause many wasted votes in the election we considered; our data are insufficient to decide.  In addition, our data cannot show a causal relationship.  In particular, while gerrymandering may cause noncompact districts, it could also be the case that states which tend to draw heavily gerrymandered districts also tend to draw noncompact districts.  However, the presence of noncompact districts does seem to generally indicate that there may be gerrymandering, and conversely states with compact districts tended to have low partisan bias.

  The individual metrics also did not perform equally well.  The area-perimeter metric did not perform particularly well, as might be expected from its simplicity: the relationship was in the right direction, but was comparatively weak.  The moment of area and moment of population preformed quite well.  Oddly, moment of area in fact did better than moment of population; we conjecture that this is due to the difficulty of properly normalizing this metric for districts of different sizes and populations.

  Metrics that use the convex hull of a district are based on the observation that problematic districts are likely to have a section ``cut out'', jut out into a neighboring district, or stretch out in a long, narrow, curved line. Like the perimeter metric, they disadvantage districts with jagged boundaries, although less heavily. The purely geometric version, like other compactness metrics, penalizes states that are non-compact. It particularly affects states that are on a coast and have islands, because a large offshore area gets counted in the convex hull. For example, California's 14th district gets an especially bad score because it combines a small, densely populated area with a group of islands several miles offshore. However, the population-based version does not have these issues. In cases where a state has complex boundaries with other states, the population outside of the state can simply be ignored. In cases where a state has islands, the offshore area has no population so it has no effect.  In addition, these metrics tend to penalize even fairly reasonable suburban districts that wrap around a city, since the convex hull often contains much of the city.

@@

  %TODO: write this section
